Integer-N Frequency Synthesizer Designed for Settling Time Compliance

### 2025-05-23 CaseStudies/Example3_2_LimitsoftheTheorySoFar/Report/article/article.qmd
Studying

@@

@@

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--', label='Frequência de corte (-3 dB)')
plt.axhline(y=-3, color='gray', linestyle=':', label='-3 dB')
plt.scatter(cutoff_freq_ang, cutoff_mag, color='red')  # Marcar o ponto

@@

plt.title('Resposta em Frequência - Diagrama de Bode')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--')
plt.ylabel('Phase (graus)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

>>>>

F_sys = signal.TransferFunction(F_num, F_den)

@@

# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = F_sys.num[0][0] if isinstance(F_sys.num[0], (list, tuple)) else F_sys.num
den = F_sys.den[0][0] if isinstance(F_sys.den[0], (list, tuple)) else F_sys.den

# Exibe como equação no Jupyter
display(Markdown(f"$$F(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-transfer-function-loop-filter}}"))

@@

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--', label='Cutoff Frequency (-3 dB)')
plt.axhline(y=-3, color='gray', linestyle=':', label='-3 dB')
plt.scatter(cutoff_freq_ang, cutoff_mag, color='red')  # Marcar o ponto

@@

plt.title('Frequency Response - Bode Plot')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--')
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

# Gerar a resposta em frequência
w, mag, phase = signal.bode(F_sys)

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')

plt.title('Bode Plot - Loop filter')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

### 2025-05-23 CaseStudies/Example3_3_IntegerNSynthesizerDesignedforSettlingTime/article/article.qmd
Studying

>>>>

# Parameters

## Maximum Phase Error Instant

Given the value of $\omega_n$, the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

$$
t_m=\frac{1}{\omega_n\sqrt{1-\zeta^2}}\tan^{-1}\left(\frac{\sqrt{1-\zeta^2}}{\zeta}\right)
$$ {#eq-instant-maximum-frequency-step}

t_m=1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

[Eq. @eq-instant-maximum-frequency-step] represents the instant at which [Eq. @eq-inverse-laplace-phase-error] reaches its maximum value.

## Maximum Normalized Phase Error

First of all, instead of using the [Eq. @eq-inverse-laplace-phase-error] and apply $t_m$, it's calculated the maximum normalized phase error, which is the same equation normalized by the input signal ($\Delta\omega$) and the natural frequency ($\omega_n$). So the maximum normalized phase error, $\theta_{e_\text{m,n}}$, is given by:

$$
\begin{split}
\theta_{e_\text{m,n}}&=\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t_m\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt_m}
\end{split}
$$

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = 2 * math.pi
```

And applying $t_m$ to this equation, the maximum normalized phase error in this case is approximately `{python} round(theta_e_m_n, 2)`.

## Maximum Allowed Frequency Step

This normalized value can be used in the following non-normalized equation for the phase error:

$$
\begin{split}
\theta_e(t)&=\frac{\Delta\omega}{\omega_n}\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt}
\end{split}
$$

The goal is to find the maximum input frequency step, $\Delta\omega_m$, such that the maximum phase error does not exceed $\theta_{e_m}=2\pi$. Using the normalized result:

$$
\begin{split}
\theta_{e_m}&=\frac{\Delta\omega_m}{\omega_n}\theta_{e_\text{m,n}}\implies\\
\Delta\omega_m&=\frac{\theta_{e_m}\omega_n}{\theta_{e_\text{m,n}}}
\end{split}
$$

step_m_ang=theta_e_m*omega_n/theta_e_m_n
step_m_lin = step_m_ang/2/math.pi
```

Substituting the calculated values yields:

- $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s
- $\Delta f_m=$ `{python} format(step_m_lin)`Hz

Hence, the maximum input frequency step that still allows theoretical predictability is `{python} format(step_m_lin)`Hz. If the input frequency step exceeds this limit, the PLL may lose lock and exhibit cycle slipping (more details are provided in the next section).

## Settling Time Estimation

The settling time, $t_s$, required for the system to reach a steady state after a frequency step can be approximated by:

t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta*omega_n)*(-1)
```

This results in a settling time of $t_s=$ `{python} format(t_s)`s.
